---
output: html_document
---

#Reproducible Research Coursera Class Peer Assessment 1

##Load required R packages and the dataset
```{r, eval=TRUE, echo=TRUE, results="asis"}
library(dplyr)
library(ggplot2)
library(grid)
library(stats)

setwd("H:/Public/Coursera/Reproducible Research")
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "./courseproject2.zip")
unzip("courseproject2.zip")
AMDdata <- read.csv("activity.csv")
AMDdata$date <- as.POSIXct(AMDdata$date, format="%Y-%m-%d")
print("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip")
summary(AMDdata)
```

##Histogram and Summary Information for the Number of Steps Taken per Day
###Histogram of Number of Steps Activity Patterns
```{r, eval=TRUE, echo=TRUE, results="asis"}
hist(AMDdata$steps, main="Number of Daily Steps Taken by an Individual")

### Summary Statistics on the Number of Steps Taken by an Individual per Day
print(paste("Mean Number of Steps = ", round(mean(AMDdata$steps, na.rm=T), digits=2)))
print(paste("Median Number of Steps = ", median(AMDdata$steps, na.rm=T)))
```

##Average Daily Activity Patterns: 5 Minute Intervals and Number of Steps
```{r, eval=TRUE, echo=TRUE, results="asis"}
meandata <- group_by(AMDdata, interval) %>% summarize(meansteps=mean(steps, na.rm=T)) %>% ungroup()
meandata$interval <- as.numeric(meandata$interval)
ggplot(data=meandata, aes(x=interval, y=meansteps)) + 
    geom_line(color="blue") +
    theme(axis.text.x = element_text(angle = 90, vjust=0.25, hjust = 1), text =            element_text(size=8), axis.title.y=element_text(vjust=1), plot.margin = unit(c(1,1.5,1,1), "cm")) +
    ggtitle("Mean Daily Steps Taken by an Individual")

meandata <- meandata[order(-meandata$meansteps), ]
print(paste("5 Minute Interval with Maximum Number of Steps identified:", meandata[1, 1]))
```

#Imputing missing values
```{r, eval=TRUE, echo=TRUE, results="asis"}
##Calcualting the Number of Missing Values
numNAs <- nrow(AMDdata) - nrow(na.omit(AMDdata))
print(paste("Number of Missing Values in the active dataset = ", numNAs))

##Replacing Missing Values with the Daily Mean Number of Steps
AMDdataNA <- group_by(AMDdata, date) %>% mutate(meansteps=mean(steps, na.rm=T)) %>% ungroup()
AMDdataNA <- mutate(AMDdataNA, TotalMeanSteps=mean(steps, na.rm=T))
AMDdataNA <- mutate(AMDdataNA, steps_sub_NA = ifelse(is.na(steps) & is.na(meansteps), TotalMeanSteps, ifelse(is.na(steps) & !is.na(meansteps), meansteps, steps)))
```

###Histogram of Number of Steps Activity Patterns
```{r, eval=TRUE, echo=TRUE, results="asis"}
hist(AMDdataNA$steps_sub_NA, main="Mean Daily Steps (NA Replaced by Daily/Total Means)")

### Summary Statistics on the Number of Steps Taken by an Individual per Day
### Missing Values were Replaced with Total (no reported values in a day) or Daily Means (at least one reported value in a day)
print(paste("Mean Number of Steps = ", round(mean(AMDdataNA$steps_sub_NA, na.rm=T), digits=2)))
print(paste("Median Number of Steps = ", round(median(AMDdataNA$steps_sub_NA, na.rm=T), digits=2)))
```

#Comparison of Weekend and Weekday Steps
```{r, eval=TRUE, echo=TRUE, results="asis"}
##Assign Weekday or Weekend to Each Date
AMDdataNA <- mutate(AMDdataNA, weekdays = weekdays(date))
AMDdataNA <- mutate(AMDdataNA, workweekornot = ifelse(weekdays %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
AMDdataNA <- group_by(AMDdataNA, interval, weekdays) %>% mutate(steps_sub_NA_meanby_weekday = mean(steps_sub_NA, na.rm=T)) %>% ungroup()

summary(AMDdataNA)

##Chart Comparing Weekend and Weekday Steps
ggplot(data=AMDdataNA, aes(x=interval, y=steps_sub_NA_meanby_weekday)) +
  geom_line(color="blue") +
  facet_wrap(~workweekornot, ncol=1) +
  ggtitle("Mean Daily Steps by Day of Week and Interval, NAs Imputed")
```